What We Eval in the Shadows

### 2021-08-12 paper/main.tex
Fix page 7
Remove invokeRestartInteractively
Remove use of eval from 5.3

@@

Our pipeline runs \CranRunnableScripts scripts extracted from \CranPackages
packages. Any run that does not call \eval is discarded, leaving us with
\packageNbruns runs. In these runs, \packageAllcalls \eval calls were recorded
originating from \PkgHitEvalCallSites sites. The runs exercised
\PkgHitEvalCallSitesAvgRatio of sites, a ratio similar to the package code
coverage metric which is \PkgCodeCoverage. The reasons some sites are not
exercised can be chalked down to incomplete tests and analysis failures
(\PkgFailedProgramsRatio of the runs crashed or timed out).
Fig.~\ref{fig:traced-eval-callsites} plots exercised site with respect to all
sites for each program; as can be seen coverage is unequal.
Table~\ref{tab:freq} summarizes the frequency of calls, call counts on the left and
number of packages on the right. There are \packageFewcalls packages with fewer
than 100 calls, and \packageManycalls packages with more than 1,000 calls.
\packageMaxcallspack calls \eval \packageMaxcalls times and thus accounts for
over half of our observations.
Table~\ref{tab:variantseval} summarizes use of variants, the first row has sites
(\emph{static}), the second has sites encountered during analysis
(\emph{exercised}) and the last gives calls (\emph{invocations}). Most sites and
calls go to \eval itself, \k{eval.parent} is rare, and both \k{evalq} and
\k{local} are barely used at all.

@@

Table~\ref{tab:provenance} summarizes provenance. Some sites are invoked with
different provenances, they are counted multiple times. Strings correlate with
dynamic code loading as done by \k{source} and \k{sys.source}. Few calls
(\packageNbParseFromFileSites in total) consume the result of calling \k{parse}
on a file. Most build strings programmatically. We identified one function
\k{invokeRestartInteractively} that prompts the user for input and \evals it.

\paragraph{Discussion}
The main use case for \eval involves using \k{match.call}, \k{substitute}, or
\k{formals} to access and transform arguments of a function. Constructing
expressions from strings is also a common use case. The constructed category
shows that most \evals comes from existing code that was modified by the
programmer before invoking \eval. Both constructed and reflection categories
roughly correspond to meta-programming. Some use cases could likely be replaced
by macros if the R designers could be convinced to overcome their distaste for
those. Another use of \eval is to directly build symbols to evaluate them in
another environment.

>>>>

Our pipeline runs \CranRunnableScripts scripts extracted from \CranPackages
packages. Any run that does not call \eval is discarded, leaving us with
\packageNbruns runs. In these runs, \packageAllcalls \eval calls were recorded
originating from \PkgHitEvalCallSites sites. The runs exercised
\PkgHitEvalCallSitesAvgRatio of sites, a ratio similar to the package code
coverage metric which is \PkgCodeCoverage. The reasons some sites are not
exercised can be chalked down to incomplete tests and analysis failures
(\PkgFailedProgramsRatio of the runs crashed or timed out).
Fig.~\ref{fig:traced-eval-callsites} plots exercised site with respect to all
sites for each program; as can be seen coverage is unequal.
Table~\ref{tab:freq} summarizes the frequency of calls, call counts on the left and
number of packages on the right. There are \packageFewcalls packages with fewer
than 100 calls, and \packageManycalls packages with more than 1,000 calls.
\packageMaxcallspack calls \eval \packageMaxcalls times and thus accounts for
over half of our observations.

@@

Table~\ref{tab:variantseval} summarizes use of variants, the first row has sites
(\emph{static}), the second has sites encountered during analysis
(\emph{exercised}) and the last gives calls (\emph{invocations}). Most sites and
calls go to \eval itself, \k{eval.parent} is rare, and both \k{evalq} and
\k{local} are barely used at all.

@@

Table~\ref{tab:provenance} summarizes provenance. Some sites are invoked with
different provenances, they are counted multiple times. Strings correlate with
dynamic code loading as done by \k{source} and \k{sys.source}. Few calls
(\packageNbParseFromFileSites in total) consume the result of calling \k{parse}
on a file. Most build strings programmatically.

\paragraph{Discussion}
The main use case for \eval involves using \k{match.call}, \k{substitute}, or
\k{formals} to access and transform arguments of a function. Constructing
expressions from strings is also a common use case. The constructed category
shows that most \evals come from existing code that was modified by the
programmer before invoking \eval. Both constructed and reflection categories
roughly correspond to meta-programming. Some use cases could likely be replaced
by macros if the R designers could be convinced to overcome their distaste for
those.
